Distributional Regression Forests for Probabilistic Precipitation Forecasting in Complex Terrain

### 2018-03-21 pkg/disttree/inst/draft/draft.Rnw
revised Section 2
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/partykit@2065 255f4249-2263-4417-b1ea-8498779e27a7

Former-commit-id: 07729d4692a600b75d13731454cad8d58097015e

\Abstract{
To obtain a probabilistic model for a dependent variable based on some set of
explanatory variables, a distributional approach is often adopted where the
parameter(s) of the distribution are linked to regressors. In many classical
models this often only captures the location/expectation of the distribution but
over the last decade there has been increasing interest in distributional
regression approaches modeling all parameters including location, scale, and
shape. Notably, so-called non-homogenous Gaussian regression models both
mean and variance of a Gaussian response variable and is particularly
popular in weather forecasting and, more generally, the GAMLSS framework
allows to establish generalized additive models for location, scale, and shape
with smooth linear or nonlinear effects.
%
However, in situations where variable selection is required and/or there are
non-smooth dependencies or interactions (especially unknown or of high-order),
it is challenging to establish a good GAMLSS. A natural alternative in these
situations would be the application of regression trees or random forests but,
so far, no general distributional framework is available for these. Therefore,
a framework for distributional trees and forests is proposed that blends
regression trees and random forests with classical distributions from the GAMLSS
framework as well as their censored or truncated counterparts.
%
To illustrate these novel approaches in practice, they are employed to obtain
probabilistic precipitation forecasts at numerous locations in a mountainous
region (Tyrol, Austria) based on a very large number of numerical weather
prediction quantities. Comparisons with GAMLSS using either a specification
based on prior meteorological knowledge or a computationally more demanding
boosting approach show that distributional random forests automatically
select variables and interactions, performing on par or often even better than
the GAMLSS approaches.
}
\Keywords{parametric models, regression trees, random forests, recursive partitioning, probabilistic forecasting, distributional regression}

@@

\begin{figure}[t!]
\centering
\setkeys{Gin}{width=0.8\textwidth}
<<rain_Axams_pred, fig=TRUE, echo=FALSE, width=6.5, height=4.5>>=
plot(x = x, y = y1, type = "l", col = "red", 
     main = paste0("July ", pday), ylab = "Density", 
     xlab = expression(Total~precipitation~"["~mm^(1/1.6)~"/"~"24h"~"]"),
     ylim = c(0,max(y1, y2, y3, y4, pm1, pm2, pm3, pm4) + 0.01),
     xlim = c(-1.5,8))

@@

With the specification of the distribution family and its log-likelihood 
function the task of fitting a distributional model turns into the task 
of estimating the distribution parameter~$\bm{\theta}$. This is commonly done by
maximum likelihood (ML) based on the learning sample
$\{y_i\}_{i = 1, \dots, n}$ of the response variable $Y$. The maximum
likelihood estimator (MLE) $\bm{\hat \theta}$ is given by
$$ \label{eq:mle}
\bm{\hat \theta} = \argmax_{\bm{\theta} \in \bold{\Theta}} \sum_{i=1}^n \ell(\bm{\theta}; y_i).
$$
Equivalently, this can be defined based on the corresponding first-order
conditions
$$ \label{eq:foc}
\sum_{i = 1}^n s(\bm{\hat \theta}, y_i) = 0,
$$
where $s(\bm{\theta}, y_i)$ is the associated score function
$$ \label{eq:score}
s(\bm{\theta}, y_i) = \frac{\partial \ell}{\partial \bm{\theta}}(\bm{\theta}; y_i).
$$ 
The latter is subsequently employed as a general goodness-of-fit measure
for individual observations $y_i$ and parameters $\bm{\theta}$.

@@

Typically, a single global model $\mathcal{D}(Y, \bm{\theta})$ is not sufficient
for reasonably representing the data and covariates $\bold{Z} = Z_1, \dots, Z_m \in \mathcal{Z}$ are
employed to capture differences in the distribution parameters $\bm{\theta}$.
In weather forecasting, these covariates typically include the output
from numerical weather prediction systems and/or lagged weather observations.

To incorporate the covariates into the distributional model, GAMLSS considers
them as regressors so that an additive predictor
$g_j(\theta_j) = f_{j,1}(\bold{Z}) + f_{j,2}(\bold{Z}) + \dots$
with link function $g_j(\cdot)$ is used for every parameter $\theta_j$
($j = 1, \dots, k$) based on smooth terms $f_{j,k}$ such as
nonlinear effects, spatial effects, random coefficients, or interaction surfaces
\citep{Klein+Kneib+Lang:2015}. However, this requires specifying the
additive terms and their functional forms in advance which is often difficult in
practice, especially if the number of covariates $m$ is large.

Regression trees generally take a different approach for automatically
including covariates in a data-driven way that also allows for abrupt
changes, non-additive effects, and interactions. In the context of
distributional models the goal is to partition the covariate space
$\mathcal{Z}$ recursively into disjoint segments so that a homogenous
distributional model for the response $Y$ can be found with segment-specific
parameters. More specifically, the $B$ disjoint segments $\mathcal{B}_b$
($b = 1, \dots, B$) partition the covariate space
$$ \label{eq:partition}
\mathcal{Z} = \dot{\bigcup\limits_{b = 1, \ldots, \textit{B}}} \mathcal{B}_b,
$$
and a local distributional model $\mathcal{D}(Y, \bm{theta}^(b))$
(i.e., with segment-specific parameters $\bm{theta}^(b)$) is fitted to the
response $Y$ in that segment.

To find the segments $\mathcal{B}_b$ that are (approximately) homogenous
with respect to the distributional model with given parameters, the idea
is to use a gradient-based recursive partitioning approach. In a given
subsample of the learning data this fits the model by ML (see
Equation~\ref{eq:mle}) and then assess the goodness of fit by looking at the
scores $s(\bm{\hat \theta}; y_i)$ (see Equation~\ref{eq:score}).

\fixme{still more to do}

These values express how well the model fits the data regarding each 
parameter. Ideally they should fluctuate randomly around zero, 
similar to residuals in ordinary least squares estimation, but in 
this case there is a score value for each pair of estimated parameter 
$\hat{\theta}_j \in (\hat{\theta}_1, \ldots, \hat{\theta}_k) = \bold{\hat{\theta}}$ 
and observation $y_j \in \{y_i\}_{i = 1,\ldots, n}$. This enables the 
model to be sensitive to changes in each of the parameters which is one 
of the main advantages of distributional trees over other tree models 
considering only the mean of the distribution.

Statistical tests are applied to check whether there is a significant 
dependency between the scores and each of the partitioning variables 
$Z_l \in \bold{Z} = (Z_1, \ldots, Z_m)$. If so, the data set is split 
into subgroups based on the values of the partitioning variable that 
shows the highest dependency. The exact way of determining how and where 
to split varies across different tree algorithms. For distributional trees 
one possible choice is a permutation test based algorithm which is explained 
in detail in the appendix.

After splitting the data the next step is to 
fit a local distributional model in each of the subgroups. Repeating this 
procedure of fitting a model, applying statistical tests to evaluate parameter 
instability and splitting the data set depending on the test results leads 
to a tree structure. In that way the learning data set is separated into 
\textit{B} disjoint segments such that
$$
\mathcal{Z} = \dot{\bigcup\limits_{b = 1, \ldots, \textit{B}}} \mathcal{B}_b
$$
where each segment is represented by a terminal node of the tree.

Summed up, the steps of building a distributional tree are:
\begin{enumerate}
\item Specify a distribution family with log-likelihood function $\ell(\theta; y)$.
\item Estimate $\bold{\hat{\theta}}$ by solving $\sum_{i=1}^n s(\hat{\theta}, y_i) = 0$.
\item Test for associations or instabilities of the scores $s(\bold{\hat{\theta}}, y_i)$ 
and each partitioning variable~$Z_l$.
\item Split the sample along the partitioning variable with the strongest association 
or instability. Choose the breakpoint with the highest improvement in log-likelihood.
\item Repeat steps 2--4 recursively in the subgroups until some stopping criterion is reached.

\end{enumerate}
The stopping criterion in step 5 can for example be a minimum number of 
observations in a node or a p-value for the statistical tests applied in step 3.
%%% <TH> we need a little more detail here or a reference to a better
%%% description of mob/ctree OR appendix</TH>

Executing these steps results in a stepwise parametric model where a complete 
distribution is fit in each node of the tree.

For distributional trees a prediction for a (possibly new) set of covariates 
$\bold{z} = (z_1, \ldots, z_m)$ is made with the following strategy: 
All observations of the learning data set $\{(y_i,\bold{z}_i)\}_{i=1\ldots,n}$ 
which end up in the same terminal node of the tree as $\bold{z}$ are selected 
and a distributional model is fit on this subset. Selecting these observations
is done by including a weight vector 
$\bold{w}^{\text{tree}} = (w^{\text{tree}}_1, \ldots, w^{\text{tree}}_n)$ 
in the fitting process where
$$
w^{\text{tree}}_i(\bold{z}) = \sum_{b=1}^B I((\bold{z}_i \in \mathcal{B}_b) \land (\bold{z} \in \mathcal{B}_b)).
$$
The predicted distribution for $\bold{z}$ is then fully 
specified by the estimated parameter $\hat{\theta}(\bold{z})$ where
$$
\hat{\theta}(\bold{z}) = \argmax_{\theta \in \Theta} \sum_{i=1}^n w^{\text{tree}}_i(\bold{z}) \cdot \ell(\theta; y_i).
$$

While the clear structure of such a tree model is advantageous in 
interpreting and analyzing it, the abrupt changes are often too rough 
and impose steps on the model even if the true underlying effect is 
smooth. In these situations bagging can solve the problem by combining 
an ensemble of trees such that wrongly detected abrupt changes of the 
model are turned into almost smooth transitions. In that way, all kinds 
of effects can be modeled, which is also one the of the main advantages 
of forest models taking bagging one step further as the variance is 
reduced by sampling the covariates considered for the splitting process 
in each node.

%%% <TH> we need to be much more precise here: random forests is Breiman's and Cutler's
%%% random forest. Maybe it is better to not mention this method at any time. </TH>
Common forest models such as the original forest technique by 
\cite{Breiman+Cutler:2004} consist of trees only considering the mean of the 
distribution of the response variable. For that reason, changes in other 
parameters can only be detected if they are connected to changes in the 
location parameter, e.g. if they depend on the location parameter. 
As explained before, a distributional tree considers scores for all 
distribution parameters. Therefore, a forest consisting of 
distributional trees can also capture separate effects in each of the 
distribution parameters. For that reason distributional forests are
now going to be introduced.

\subsection{Distributional forest}
\label{distforest}
A distributional forest consists of an ensemble of $T$ distributional trees. 
Each of these trees is built on a different data set which can either be a bootstrap 
or a subset of the original data set. Moreover, in each node a subset of the 
partitioning variables is chosen randomly as candidates for splitting variables. 
In that way the correlation among the resulting trees is reduced and therefore also 
the variance of the model (\citealt{Breiman:2001}).

Forest models differ mainly in the way how the trees are combined to make predictions. 
Following the strategy explained and proposed by \cite{Hothorn+Zeileis:2017} we 
interpret random forests as adaptive local likelihood estimators as distributional 
models are fit locally in subgroups while the term ``adaptive'' refers to the 
dependence on the splitting variables $\bold{Z}$ and the resulting subgroups 
and weighing schemes. In this case the forest itself is only used to evaluate 
neighborhood relationships among the observations which can then be included 
in the fitting process. In transformation forests by \cite{Hothorn+Zeileis:2017} 
so-called ``nearest neighbor weights'' are calculated as a measure of distance 
between observations. This approach has also already been applied in several 
random forest-type methods such as in ``adaptive nearest neighbors'' by 
\cite{Lin+Jeon:2006}, in ``quantile regression forests''  by  \cite{Meinshausen:2006} 
and in ``bagging survival trees'' by \cite{Hothorn+Lausen+Benner:2004}. 
The idea of nearest neighbor weights is to measure similarity of a pair of 
observations by counting the number of trees in which it is asigned to the 
same terminal node.
%Similar observations have a high probability of ending up in the same terminal 
%node whereas this probability is low for quite different observations. 
%although the first random forest-type algorithm for the estimation of 
%conditional distribution functions was published more than a decade ago 
%(bagging survival trees, Hothorn et al. 2004). 
In particular, for a (possibly new) set of covariates 
$\bold{z} = (z_1, \ldots, z_m) $  such a set of weights 
$\bold{w}^{\text{forest}}~=~(w^{\text{forest}}_1, \ldots, w^{\text{forest}}_n)$ 
is obtained by the following procedure: For each tuple $(y_i,\bold{z}_i)$ in 
the learning data set $\{(y_i,\bold{z}_i)\}_{i=1\ldots,n}$ a weight is calculated 
by counting the number of trees in which $(y_i,\bold{z}_i)$ ends up in the same 
terminal node as $\bold{z}$:
$$
w^{\text{forest}}_i(\bold{z}) = \frac{1}{T} \sum_{t=1}^T \sum_{b=1}^{B^t} I((\bold{z}_i \in \mathcal{B}^t_b) \land (\bold{z} \in \mathcal{B}^t_b))
$$
where $T$ is the number of trees of the forest, $B^t$ is the number of terminal 
nodes of the \textit{t}-th tree and $\mathcal{B}^t_b$ represents the \textit{b}-th 
terminal node of the \textit{t}-th tree.

Therefore, contrary to distributional trees, in distributional forests the whole 
learning data set can be used to fit a distributional model and specify the 
predicted distribution for a set of covariates. Compared to the formula for
predictions made by distributional trees the vector of tree weights 
$\bold{w}^{\text{tree}}$ containing only zeros and ones is now replaced by a 
vector of integer valued forest weights.
%%% <TH> we need references for the nearest neighbor weights, copy from
%%% "Transformation Forests" </TH>
%%%<TH> we need to cite "Transformation Forests" for the interpretation of
%%% random forests as adaptive local ML </TH>
This individual set of weights 
$\bold{w}^{\text{forest}}(\bold{z}) = (w^{\text{forest}}_1(\bold{z}), \ldots, w^{\text{forest}}_n(\bold{z}))$ 
for the set of covariates $\bold{z}$ can now be included in the 
estimation process of the distribution parameters and in that way leads 
to its individual parameter vector $\hat{\theta}(\bold{z})$ where
$$
\hat{\theta}(\bold{z}) =  \argmax_{\theta \in \Theta} \sum_{i=1}^n w^{\text{forest}}_i(\bold{z}) \cdot \ell(\theta; y_i)
$$
While in a distributional tree the same distribution is specified for all 
sets of covariates being in the same node, a distributional forest predicts 
and fits a different distribution for each different set of covariates due 
to the individual weights calculated separately for each set of covariates.

As this novel methodology combines the advantages of smooth and flexible 
parametric modeling with those of tree-structured modeling it provides a 
model that can deal with a wide range of situations. Abrupt changes and 
non-additive effects can be captured by its tree structure while smooth 
effects can be modeled reasonably well by combining various tree and 
smoothing the steps imposed by a single tree. As scores are considered 
for each distribution parameter (separate) effects in different parameters 
can be captured. At the same time, also interactions are detected 
automatically by splitting the data in a way that is sensitive to changes 
in either each of the parameters or combined in various parameters.

@@

From the GEFS outputs 14 basic forecast variables are considered with up 
to 12 variations of each of these variables such as mean/maximum/minimum
or different aggregation time ranges, etc. A detailed overview is
provided in Table~\ref{covariates}, yielding 80~predictor variables in total. (Since the minimum of the ``variable downwards short wave radiation flux'' (\emph{dswrf}) over 24 hours is always 0, the ensemble mean and standard deviation of these minimal values are not considered.) \fixme{Mention that dswrf\_mean/sprd\_min is not included?}

@@

\subsection{Application for one station}
\label{Application for one station}
First of all we focus on one observation station, Axams in Tyrol. This 
station is selected due to its geographical closeness to Innsbruck, 
the capital of Tyrol. As for all other stations, daily observations and 
numerical weather predictions are available for the month of July of 
28~years from 1985 through the year 2012. In Figure~\ref{axams_onestation} 
the results of fitting a distributional forest to the first 24 years of the 
data set and predicting the total amount of precipitation for one specific 
day of each of the four successive years have been shown as a motivational 
example for a probabilistic forecast. Staying with this fitted model,
variable importance of the distributional forest model will be analyzed and its
performance will be compared to those of the prespecified and the boosted GAMLSS 
and the EMOS model based on residual QQ-plots.

@@

\section{Tree algorithm}
\label{treealgorithm}
The tree algorithm used in the applications discussed in this paper 
is going to be explained following the ideas presented in 
\cite{Seibold+Zeileis+Hothorn:2017}. For notational simplicity, the testing 
and splitting procedure is described for the root node with observations 
$\{y_i\}_{i = 1,\ldots,n}$, $n \in \mathbb{N}$. In each child node the 
corresponding subset of observations depends on the assignment made 
by the foregoing split.

\begin{itemize}
\item A distributional model $\mathcal{D}(Y, \theta)$ is fit to the set of 
observations $\{y_i\}_{i = 1,\ldots,n}$ by estimating the distribution 
parameter $\hat{\theta}$. This parameter can be a multidimensional 
vector $(\hat{\theta}_1, \ldots, \hat{\theta}_K)$, $K \in \mathbb{N}$, 
depending on the type of distribution. The estimation is done by 
maximizing the log-likelihood function $\ell(\theta; Y)$, i.e.
$$
\hat{\theta} = \argmax_{\theta \in \Theta} \sum_{i=1}^n \ell(\theta; y_i)
$$
which is equal to solving
$$
\sum_{i=1}^n s(\theta, y_i) = \sum_{i=1}^n \frac{\partial \ell}{\partial \theta}(\theta; y_i) = 0
$$
for $\theta$.

>>>>

\Abstract{
To obtain a probabilistic model for a dependent variable based on some set of
explanatory variables, a distributional approach is often adopted where the
parameter(s) of the distribution are linked to regressors. In many classical
models this only captures the location/expectation of the distribution but
over the last decade there has been increasing interest in distributional
regression approaches modeling all parameters including location, scale, and
shape. Notably, so-called non-homogenous Gaussian regression (NGR) models both
mean and variance of a Gaussian response and is particularly
popular in weather forecasting and, more generally, the GAMLSS framework
allows to establish generalized additive models for location, scale, and shape
with smooth linear or nonlinear effects.
%
However, when variable selection is required and/or there are
non-smooth dependencies or interactions (especially unknown or of high-order),
it is challenging to establish a good GAMLSS. A natural alternative in these
situations would be the application of regression trees or random forests but,
so far, no general distributional framework is available for these. Therefore,
a framework for distributional trees and forests is proposed that blends
regression trees and random forests with classical distributions from the GAMLSS
framework as well as their censored or truncated counterparts.
%
To illustrate these novel approaches in practice, they are employed to obtain
probabilistic precipitation forecasts at numerous locations in a mountainous
region (Tyrol, Austria) based on a very large number of numerical weather
prediction quantities. It is shown that distributional random forests
automatically select variables and interactions, performing on par or often even
better than GAMLSS specified either through prior meteorological knowledge or a
computationally more demanding boosting approach.
}
\Keywords{parametric models, regression trees, random forests, recursive partitioning, probabilistic forecasting, distributional regression}

@@

\begin{figure}[t!]
\centering
\setkeys{Gin}{width=0.8\textwidth}
<<rain_Axams_pred, fig=TRUE, echo=FALSE, width=6.5, height=4>>=
par(mar = c(4, 4, 3, 1.5))
plot(x = x, y = y1, type = "l", col = "red", 
     main = paste0("July ", pday), ylab = "Density", 
     xlab = expression(Total~precipitation~"["~mm^(1/1.6)~"/"~"24h"~"]"),
     ylim = c(0,max(y1, y2, y3, y4, pm1, pm2, pm3, pm4) + 0.01),
     xlim = c(-1.5,8))

@@

With the specification of the distribution family and its log-likelihood 
function the task of fitting a distributional model turns into the task 
of estimating the distribution parameter~$\bm{\theta}$. This is commonly done by
maximum likelihood (ML) based on the learning sample
$\{y_i\}_{i = 1, \dots, n}$ of the response variable $Y$. The maximum
likelihood estimator (MLE) $\bm{\hat \theta}$ is given by
$$ \label{eq:mle}
\bm{\hat \theta} = \argmax_{\bm{\theta} \in \bold{\Theta}} \sum_{i=1}^n \ell(\bm{\theta}; y_i).
$$
Equivalently, this can be defined based on the corresponding first-order
conditions
$$ \label{eq:foc}
\sum_{i = 1}^n s(\bm{\hat \theta}, y_i) = 0,
$$
where $s(\bm{\theta}, y_i)$ is the associated score function
$$ \label{eq:score}
s(\bm{\theta}, y_i) = \frac{\partial \ell}{\partial \bm{\theta}}(\bm{\theta}; y_i).
$$ 
The latter is subsequently employed as a general goodness-of-fit measure
to assess how well the distribution with parameters $\bm{\theta}$ fits
one individual observation $y_i$.

@@

Typically, a single global model $\mathcal{D}(Y, \bm{\theta})$ is not sufficient
for reasonably representing the response distribution and covariates
$\bold{Z} = Z_1, \dots, Z_m \in \mathcal{Z}$ are
employed to capture differences in the distribution parameters $\bm{\theta}$.
In weather forecasting, these covariates typically include the output
from numerical weather prediction systems and/or lagged weather observations.

To incorporate the covariates into the distributional model, GAMLSS considers
them as regressors in additive predictors
$g_j(\theta_j) = f_{j,1}(\bold{Z}) + f_{j,2}(\bold{Z}) + \dots$.
Link functions $g_j(\cdot)$ are used for every parameter $\theta_j$
($j = 1, \dots, k$) based on smooth terms $f_{j,k}$ such as
nonlinear effects, spatial effects, random coefficients, or interaction surfaces
\citep{Klein+Kneib+Lang:2015}. However, this requires specifying the
additive terms and their functional forms in advance which can be challenging in
practice, especially if the number of covariates $m$ is large.

Regression trees generally take a different approach for automatically
including covariates in a data-driven way and allowing for abrupt
changes, non-additive effects, and interactions. In the context of
distributional models the goal is to partition the covariate space
$\mathcal{Z}$ recursively into disjoint segments so that a homogenous
distributional model for the response $Y$ can be found with segment-specific
parameters. More specifically, the $B$ disjoint segments $\mathcal{B}_b$
($b = 1, \dots, B$) partition the covariate space
$$ \label{eq:partition}
\mathcal{Z} = \dot{\bigcup\limits_{b = 1, \ldots, \textit{B}}} \mathcal{B}_b,
$$
and a local distributional model $\mathcal{D}(Y, \bm{\theta}^{(b)})$
(i.e., with segment-specific parameters $\bm{\theta}^{(b)}$) is fitted to the
response $Y$ in that segment.

To find the segments $\mathcal{B}_b$ that are (approximately) homogenous
with respect to the distributional model with given parameters, the idea
is to use a gradient-based recursive-partitioning approach. In a given
subsample of the learning data this fits the model by ML (see
Equation~\ref{eq:mle}) and then assesses the goodness of fit by assessing the
corresponding scores $s(\bm{\hat \theta}; y_i)$ (see Equation~\ref{eq:score}).

In summary, distributional trees are fitted recursively via:
\begin{enumerate}
\item Estimate $\bm{\hat \theta}$ via maximum likelihood for the observations
  in the current subsample.
\item Test for associations or instabilities of the scores $s(\bm{\hat \theta}, y_i)$ 
  and $Z_{l,i}$ for each partitioning variable~$Z_l$ ($l = 1, \dots, m$).
\item Split the sample along the partitioning variable $Z_{l,i}^*$ with the
  strongest association or instability. Choose the breakpoint with the highest
  improvement in the log-likelihood or the highest discrepancy .
\item Repeat steps 1--3 recursively in the subsamples until these become too
  small or there is no significant association/instability (or some other
  stopping criterion is reached).
\end{enumerate}
%
Different inference techniques can be used for assessing the association between
scores and covariates in step~3. In the following we use the general class of
permutation tests introduced by \cite{Hothorn+Hornik+VanDeWiel:2006} which is
also the basis of conditional inference trees \citep[CTree,][]{Hothorn+Hornik+Zeileis:2006}.
Alternatively, one could use asymptotic M-fluctuation tests for parameter
instability \citep{Zeileis+Hornik:2007} as in model-based recursive partitioning
\citep[MOB,][]{Zeileis+Hothorn+Hornik:2008}. More details are provided in
Appendix~\ref{app:tree}.

For obtaining probabilistic predictions from the tree for a (possibly new) set
of covariates $\bold{z} = (z_1, \ldots, z_m)$, the observation simply has to
be ``sent down'' the tree and the corresponding segment-specific MLE has to be
obtained. This can also be understood as a weighted MLE where the weights
select those observations from the learning sample that fall into the same
segment:
$$
w^{\text{tree}}_i(\bold{z}) = \sum_{b=1}^B \mathbf{1}((\bold{z}_i \in \mathcal{B}_b) \land (\bold{z} \in \mathcal{B}_b)),
$$
where $\mathbf{1}(\cdot)$ is the indicator function. The predicted distribution
for a given $\bold{z}$ is then fully specified by the estimated parameter
$\bm{\hat \theta}(\bold{z})$ where
$$
\bm{\hat \theta}(\bold{z}) = \argmax_{\bm{\theta} \in \bm{\Theta}} \sum_{i=1}^n w^{\text{tree}}_i(\bold{z}) \cdot \ell(\bm{\theta}; y_i).
$$

\subsection{Distributional forest}
\label{distforest}

While the simple recursive structure of a tree model is easy to visualize and
interpret, the abrupt changes are often too rough, instable, and impose steps
on the model even if the true underlying effect is smooth. Hence, ensemble
methods such as bagging or random forests \citep{Breiman:2001} typically smooth
the effects, stabilize the model, and improve predictive performance.

While classical random forests \citep{Breiman:2001} grow ensembles of trees
that pick up changes in the location of the response across the covariates,
\emph{distributional forests} employ an ensemble of $T$~\emph{distributional trees}. These
pick up changes in the ``direction'' of any distribution parameter by
considering the full score vector for choosing splitting variables and split
points. Each of the distributional trees is grown on a different data set
obtained through bootstrap sampling (or subsampling) and in each node only a
random subset of the covariates $\bold{Z}$ is considered. As usual in random
forests, this reduces the correlation among the trees and stabilizes the
variance of the model.

To obtain probabilistic predictions from a distributional forest, it still
needs to be specified how to compute the parameter estimates
$\bm{\hat \theta}(\bold{z})$ for a (potentially new) set of covariates $\bold{z}$.
Following \cite{Hothorn+Zeileis:2017} we interpret random forests as adaptive
local likelihood estimators using the averaged ``nearest neighbor weights''
\citep{Lin+Jeon:2006} from the $T$~trees in the forest
$$
w^{\text{forest}}_i(\bold{z}) = \frac{1}{T} \sum_{t=1}^T \sum_{b=1}^{B^t} \mathbf{1}((\bold{z}_i \in \mathcal{B}^t_b) \land (\bold{z} \in \mathcal{B}^t_b))
$$
Thus, these $w^{\text{forest}}_i(\bold{z}) \in [0, 1]$ whereas
$w^{\text{tree}}_i(\bold{z}) \in \{0, 1\}$. Hence, weights cannot only be $0$
or $1$ but change more smoothly, giving high weight to those observations $i$
from the learning sample that co-occur in the same segment $\mathcal{B}_b^t$
as the new observation $\bold{z}$ for many of the trees $t = 1, \dots, T$.
Consequently, the parameter estimates may, in principle, change for every
observation and can be obtained by
$$
\bm{\hat \theta}(\bold{z}) =  \argmax_{\bm{\theta} \in \bm{\Theta}} \sum_{i=1}^n w^{\text{forest}}_i(\bold{z}) \cdot \ell(\bm{\theta}; y_i).
$$
In summary, this yields a parametric distributional regression model 
(through the score-based approach) that can capture both abrupt effects and
high-order interactions (through the trees) and smooth effects (through
the forest).

@@

From the GEFS outputs 14 basic forecast variables are considered with up 
to 12 variations of each of these variables such as mean/maximum/minimum
or different aggregation time ranges, etc. A detailed overview is
provided in Table~\ref{covariates}, yielding 80~predictor variables in total.
(Since the minimum of the ``variable downwards short wave radiation flux'',
\emph{dswrf}, over 24 hours is always 0, the ensemble mean and standard
deviation of these minimal values are not considered.)

@@

\subsection{Application for one station}
\label{Application for one station}

In a first step, we show a detailed comparison of the competing models
for one observation station, Axams in Tyrol (selected somewhat arbitrarily
as the closest in the data set to Innsbruck, capital of Tyrol). As for all
other stations, daily observations and  numerical weather predictions are available for the month of July of 
28~years from 1985 through the year 2012. In Figure~\ref{axams_onestation} 
the results of fitting a distributional forest to the first 24 years of the 
data set and predicting the total amount of precipitation for one specific 
day of each of the four successive years have been shown as a motivational 
example for a probabilistic forecast. Staying with this fitted model,
variable importance of the distributional forest model will be analyzed and its
performance will be compared to those of the prespecified and the boosted GAMLSS 
and the EMOS model based on residual QQ-plots.

@@

\section{Tree algorithm} \label{app:tree}
The tree algorithm used in the applications discussed in this paper 
is going to be explained following the ideas presented in 
\cite{Seibold+Zeileis+Hothorn:2017}. For notational simplicity, the testing 
and splitting procedure is described for the root node with observations 
$\{y_i\}_{i = 1,\ldots,n}$, $n \in \mathbb{N}$. In each child node the 
corresponding subset of observations depends on the assignment made 
by the foregoing split.

\begin{itemize}
\item A distributional model $\mathcal{D}(Y, \bm{\theta})$ is fit to the set of 
observations $\{y_i\}_{i = 1,\ldots,n}$ by estimating the distribution 
parameter $\bm{\hat \theta}$. This parameter can be a multidimensional 
vector $(\hat{\theta}_1, \ldots, \hat{\theta}_K)$, $K \in \mathbb{N}$, 
depending on the type of distribution. The estimation is done by 
maximizing the log-likelihood function $\ell(\bm{\theta}; Y)$, i.e.
$$
\hat{\theta} = \argmax_{\theta \in \Theta} \sum_{i=1}^n \ell(\theta; y_i)
$$
which is equal to solving
$$
\sum_{i=1}^n s(\theta, y_i) = \sum_{i=1}^n \frac{\partial \ell}{\partial \theta}(\theta; y_i) = 0
$$
for $\theta$.
